Abstract Data Type Inference

### 2015-02-28 draft.tex
Added to related work.

  Checking that efficient implementations of concurrent objects adhere to their
  abstract data types (ADTs) can be made tractable when ADTs are given by
  symbolic, logical representations. Obtaining these symbolic representations,
  however, is a complex task requiring rare expertise. In practice, concurrent
  objects are specified indirectly by possibly-abstract imperative reference
  implementations.

  In this work we demonstrate that effective symbolic ADT representations can
  be automatically generated from the executions of reference implementations.
  Our approach exploits key features of naturally-occurring ADTs: violations
  can be decomposed into a small set of representative patterns, and these
  patterns manifest in executions with few operations. By identifying certain
  algebraic properties of naturally-occurring ADTs, and exhaustively sampling
  executions up to a small number of operations, we generate concise symbolic
  ADT representations which are complete in practice.

>>>>

  Checking that efficient implementations of concurrent objects adhere to their
  abstract data types (ADTs) can be made tractable when ADTs are given by
  symbolic, logical representations. Obtaining these symbolic representations,
  however, is a complex task requiring rare expertise. In practice, concurrent
  objects are specified indirectly by imperative reference implementations.

  In this work we demonstrate that effective symbolic ADT representations can
  be automatically generated from the executions of reference implementations.
  Our approach exploits key features of naturally-occurring ADTs: violations
  can be decomposed into a small set of representative patterns, and these
  patterns manifest in executions with few operations. By identifying certain
  algebraic properties of naturally-occurring ADTs, and exhaustively sampling
  executions up to a small number of operations, we generate concise symbolic
  ADT representations which are complete in practice, enabling the application
  of efficient symbolic verification algorithms without the burden of manual
  specification.

### 2015-02-28 intro.tex
Added to related work.

Although checking observational refinement is a fundamentally hard
problem\footnote{See Section~\ref{sec:related} for further discussion.} for
arbitrary ADTs~\cite{journals/siamcomp/GibbonsK97, conf/esop/BouajjaniEEH13},
recent work demonstrates that naturally-occurring ADTs like atomic collections,
semaphores, and locks can be checked efficiently given effective symbolic ADT
representations~\cite{conf/popl/BouajjaniEEH15, journals/arxiv/BouajjaniEEH15,
conf/pldi/EmmiEH15}, e.g.,~specified in first-order logic. However, writing
accurate symbolic specifications is an expected challenge for ADT designers. In
practice, ADTs are specified indirectly as the observable behaviors of
possibly-abstract imperative \emph{reference implementations}. Mapping these
reference implementations to logical formulæ describing the same sets of
behaviors is a complex task requiring rare expertise.

In this work we demonstrate that effective symbolic ADT representations can be
generated automatically from the executions of reference implementations. Our
approach exploits two key features of concurrent-object ADTs: that violations
of each ADT can be decomposed into a small set of representative patterns, and
that these patterns manifest in executions with few operations. The first
feature allows us to represent symbolic ADTs finitely, as exclusions of
violation patterns. The second allows us to extract violation patterns from
finite enumerations of executions.

@@

>>>>

Although checking observational refinement is a fundamentally hard
problem\footnote{See Section~\ref{sec:related} for further discussion.} for
arbitrary ADTs~\cite{journals/siamcomp/GibbonsK97, conf/esop/BouajjaniEEH13},
recent work demonstrates that implementations of naturally-occurring ADTs like
atomic collections, semaphores, and locks can be checked efficiently given
effective symbolic ADT representations~\cite{conf/popl/BouajjaniEEH15,
journals/arxiv/BouajjaniEEH15, conf/pldi/EmmiEH15}, e.g.,~specified in
first-order logic. However, writing accurate symbolic specifications is an
expected challenge for ADT designers. In practice, ADTs are specified
indirectly as the observable behaviors of possibly-abstract imperative
\emph{reference implementations}. Mapping these reference implementations to
logical formulæ describing the same sets of behaviors is a complex task
requiring rare expertise.

In this work we demonstrate that effective symbolic ADT representations can be
generated automatically from the executions of reference implementations,
enabling the application of efficient symbolic refinement-checking algorithms
without the burden of writing formal specifications manually. Our approach
exploits two key features of concurrent-object ADTs: that violations of each
ADT can be decomposed into a small set of representative patterns, and that
these patterns manifest in executions with few operations. The first feature
allows us to represent symbolic ADTs finitely, as exclusions of violation
patterns. The second allows us to extract violation patterns from finite
enumerations of executions.

@@

As far as we are aware, this work is the first to suggest the automatic
generation of symbolic ADT representations for concurrent objects. By removing
the burden of writing formal ADT specifications manually, this work broadens
the scope of efficient symbolic refinement-checking algorithms to
newly-designed ADTs, apart from those few traditionally studied in the
literature.

### 2015-02-28 related.tex
Added to related work.

TODO COMPARE WITH OTHER REFINEMENT CHECKING APPROACHES AND FIGURE OUT HOW THIS
MIGHT BE RELATED TO OTHER INFERENCE ALGORITHMS

>>>>

Inference of symbolic abstract data types is a recently-pertinent problem
arising from the the emergence of efficient symbolic refinement-checking
algorithms~\cite{conf/popl/BouajjaniEEH15, conf/pldi/EmmiEH15} which require
symbolic, logical representations of ADT specifications. Though approaches
based on the explicit enumeration of execution
linearizations~\cite{journals/jpdc/WingG93, conf/pldi/BurckhardtDMT10,
conf/asplos/BurnimNS11, conf/kbse/ZhangCW13} do not require symbolic ADT
representations, and work directly with given reference implementations, these
approaches are intractable, since the number of linearizations is exponential
in execution length.

Other approaches to tractable refinement-checking are based on annotating
method bodies with \emph{linearization
points}~\cite{journals/toplas/HerlihyW90, conf/cav/AmitRRSY07,
conf/fm/LiuCLS09, conf/cav/Vafeiadis10, conf/podc/OHearnRVYY10,
conf/icse/Zhang11a, conf/oopsla/ShachamBASVY11, conf/cav/DragoiGH13,
conf/pldi/LiangF13} to reduce the otherwise exponential number of possible
linearizations to one single linearization. This approach, however, does not
apply to all implementations~\cite{journals/toplas/HerlihyW90}, and
applications of this approach often rely on manual annotation of the
implementation source code. Furthermore, this approach does not admit
conclusive evidence of a refinement violation in case of failure.

The idea of decomposing ADT violations into a finite set of patterns has been
proposed for atomic collections~\cite{conf/tacas/AbdullaHHJR13,
conf/concur/HenzingerSV13, conf/popl/DoddsHK15, conf/popl/BouajjaniEEH15}, and
generalized to ADTs which can be expressed with a particular form of recursive
definition~\cite{journals/arxiv/BouajjaniEEH15}. These works suggest patterns
which directly recognize violations in \emph{concurrent} executions,
effectively reducing observational refinement to the verification of classical
temporal-logic properties. Some of these patterns are expressible only as
\emph{infinite}, though regular, languages. On the contrary, in this work we
suggest patterns expressed as \emph{finite} languages, which recognize
violations in ADT \emph{kernels} — i.e.,~the sequential executions of atomic
objects, or the “concurrency-aware” executions of “concurrency-aware”
specifications~\cite{conf/podc/HemedR14}. Though by themselves these patterns
are incomplete in recognizing violations in concurrent executions, they are
complete when considered as negative conditions on the linearizations of
concurrent executions, effectively reducing verification of a single execution
to satisfiability checking~\cite{conf/pldi/EmmiEH15}.
